### 2023-03-03 Manuscript/Draft1.Rmd
manuscript update

We compare abundance estimates from stratified mean calculations to those derived from the Vector Autoregressive Spatio-Temporal (VAST) model. The stratified mean is a design-based approach that calculates the geometric mean catch per tow and has traditionally been used with stratified random sample designs. VAST is a spatial delta-generalized linear mixed model that estimates both abundance (biomass) and probability of occurrence (presence/absence) [@thorson2019guidance]. If desired, VAST also allows users to include covariate data to better inform the model. Covariates can be static (eg. habitat preferences) or dynamics (eg. temperature). We explore whether including environmental predictors can help inform models and provide better abundance estimates, which is particularly relevant as climate change progresses. The stratified mean calculations are straightforward and quick, while VAST models require numerous user inputs and can take on the order of hours to complete. We compare the yearly estimated of abundance obtained from the stratified mean to estimates obtained from the Vector-Autoregressive Spatio-temporal (VAST) model.

We follow the advice given in [@thorson2019guidance] to build VAST models to estimate biomass on Georges Bank using stratified mean samples from our model output. In addition to exploring different link functions and assumed distributions, our VAST model-building process included testing the impact of including spatial and/or spatio-temporal variation in our models, considering varying number of knots in our mesh, and testing different forms of temporal correlation. We also carried out the same process both including covariates in our model as well as running models without covariate information. We considered covariates in the form of dynamic temperature values and/or static habitat values from our population model. When using covariates we ultimately decided to provide the most information to the model by including both covariates for both linear predictors. Since we know the true population values in our models we calculate the absolute error of each VAST estimate to compare between potential settings. Through this process, and in consultation with the VAST package creator, we determined setting that allowed VAST models to converge for all of our scenarios while also providing the lowest absolute error values. Settings for our VAST models can be seen in Table ???.

Our goal is to determine indexing approaches and settings that are robust to future environmental conditions and resulting spatial biomass patterns. An underlying assumption in all indexing methods is that individual random samples combine to accurately represent true abundance by a) containing a low enough noise level in the samples to allow for a discernible pattern and b) sampling all strata in which the population exists. These assumptions can be questioned given enough noise in the sampling process **cite?** and/or climate change causing a population to move into previously uninhabited strata. To simulate the impact of noise, we compare indexing estimates after adding noise to our samples versus those using the true sampling values. **BL: Help with correct notation for adding noise**. We simulate the effect of populations moving into new habitat by comparing indexing estimates using samples from all strata inhabited by each species on Georges bank [[Bank]] to those that only include a subset of the full spatial domain for each species. We chose strata to exclude for each species by reviewing how spatial preferences evolved in our increasing temperature scenarios and removing strata that each species either shifted into, or away from. Tabl XXX [[needs to be added]] lists all strata inhabited by each species, those that are removed from certain calculations, and the explanation of why these strata were removed.

**BL: I tried to provid emore detail above and also added another table**
[[may need to spell the aspect of reduced spatial domain a bit more because it might be confusing to readers who are expected us to just add areas around the current ones instead of reducing the strata]]

@@

The goal of our project was to analyze how well contemporary population indexing methods can track population trends under a variety of conditions, as depicted in Table \@ref(tab:scenarios) Historically, Atlantic Cod has seen significant decline over the last XXX years while Haddock has increased in abundance in recent year [[can cite the 2022 management track assessments, see https://apps-nefsc.fisheries.noaa.gov/saw/sari.php for when the document becomes available]] **cite**. For this reason we compare indexing estimates using stratified random samples from decreasing population scenarios for Cod (see Table ???) and increasing population scenarios for Haddock (see Table ???). To provide a comprehensive analysis of population indexing methods we consider all possible scenario combinations for Yellowtail Flounder (see Table ???). The specific population trends used in our analyses can be see in Figure \@ref(fig:pop-scenarios)

@@

General themes that exist in Tables X, Y and Z are that VAST estimates provide lower errors relative to those derived from the stratified mean, with VAST models that include covariate information providing the lowest overall errors. We also see individual cases where the stratified mean produced the lowest absolute error and instances where including covarites in VAST models actually increase the absolute error. When we reduce the number of strata that are included in indexing calculations to simulation species shifting into new territory, we typically see an increase in absolute error (as expected), though there are some scenarios where the impact is minimal.

In the Cod results, when using reduced strata adding covariates produces worst VAST results (though still better than stratified mean). VAST without covariates much worse than stratified mean in fall with increasing temperature and all strata, but adding covariates corrects this.

For Haddock, VAST has a particularly hard time in spring regularly producing larger errors than the stratified mean with added covariates only improving to the level of the stratified mean. VAST shows improved results in fall relative to the stratified mean with added covaraites producing extremely low errors in some cases.

@@

@@

@@

tabl <- "
|	Parameter	|	Description | Input/Value
|:-	|:------- |:----
| *ObsModel* | Link function and assumed distribution | c(10,2)
|FieldCOnfig | Specified spatial and/or spatio-temporal variation in predictors| c(Omega1=0, Epsilon1=0, Omega2=1, Epsilon2=1)
| RhoConfig | Specifying whether intercepts or spatio-temporal variation is structured among time intervals | c(Beta1=3, Beta2=3, Epsilon1=0, Epsilon2=4)
| X1_formula | Right-sided formula affecting the 1st linear predictor  | X1_formula = ~ poly(Temp, degree=2 )
| X2_formula | Right-sided formula affecting the 2nd linear predictor | X2_formula = ~ poly(Temp, degree=2 ) + poly(Habitat, degree=2 )

>>>>

We compare yearly abundance estimates obtained from the stratified mean to estimates obtained from the Vector-Autoregressive Spatio-temporal (VAST) model. The stratified mean is a design-based approach that calculates the geometric mean catch per tow and has traditionally been used with stratified random sample designs. VAST is a spatial delta-generalized linear mixed model that estimates both abundance (biomass) and probability of occurrence (presence/absence) [@thorson2019guidance]. If desired, VAST also allows users to include covariate data to better inform the model. Covariates can be static (eg. habitat preferences) or dynamics (eg. temperature). We explore whether including environmental predictors can help inform models and provide better abundance estimates, which is particularly relevant as climate change progresses. The stratified mean calculations are straightforward and quick, while VAST models require numerous user inputs and can take on the order of hours to complete.

We follow the advice given in [@thorson2019guidance] to build VAST models to estimate biomass on Georges Bank using stratified random samples from our model output. In addition to exploring different link functions and assumed distributions, our VAST model-building process involved testing the impact of including spatial and/or spatio-temporal variation in our models, considering varying number of knots in our mesh, and testing different forms of temporal correlation. We also carried out the same process running models without covariate information as well as including covariates in our model. We considered covariates in the form of dynamic temperature values and/or static habitat values ($Hab_{J.s}$) from our population model. When using covariates we ultimately decided to provide the most information to the model by including both temperature and habitat covariates for both linear predictors. Knowing the true population values in our models allowed us to calculate the absolute error of each VAST estimate to compare between potential settings. Through this process, and in consultation with the VAST package creator, we ultimately compared the performance of two sets of settings in our VAST models, which can be seen in Table\@ref(tab:VASTsettings).

Our goal is to determine indexing approaches and settings that are robust to future environmental conditions and resulting spatial biomass patterns. An underlying assumption in all indexing methods is that individual random samples combine to accurately represent true abundance by a) containing a low enough noise level in the samples to allow for a discernible pattern and b) sampling all strata in which the population exists. These assumptions can be questioned given enough noise in the sampling process **cite?** and/or climate change causing a population to move into previously uninhabited strata. To simulate the impact of noise, we compare indexing estimates after adding noise to our samples versus those using the true sampling values. **BL: Help with correct notation for adding noise**. We simulate the effect of populations moving into new habitat by comparing indexing estimates using samples from all strata inhabited by each species on Georges Bank to those that only include a subset of the full spatial domain for each species. We chose strata to exclude for each species by reviewing how spatial preferences evolved in our increasing temperature scenarios and removing strata that each species either shifted into, or away from. Tabl XXX [[needs to be added]] lists all strata inhabited by each species, those that are removed from certain calculations, and the explanation of why these strata were removed.

@@

The goal of our project was to analyze the ability of contemporary population indexing methods to track population trends under a variety of conditions, as depicted in Table \@ref(tab:scenarios). Historically, Atlantic Cod has seen significant decline over the last XXX years while Haddock has increased in abundance in recent year [[can cite the 2022 management track assessments, see https://apps-nefsc.fisheries.noaa.gov/saw/sari.php for when the document becomes available]] **cite**. For this reason we compare indexing estimates using stratified random samples from decreasing population scenarios for Cod and increasing population scenarios for Haddock. To provide a comprehensive analysis of population indexing methods we consider all possible scenario combinations for Yellowtail Flounder. The specific population trends used in our analyses can be see in Figure \@ref(fig:pop-scenarios).

@@

Tables \@ref(tab:YTresultsallstrata), \@ref(tab:YTresultsreducedstrata), \@ref(tab:CodERROR), and \@ref(tab:HadERROR) contain the results of our comparison between the absolute error of abundance estimates and model output. General themes are that VAST estimates provide lower errors relative to those derived from the stratified mean, with VAST models that include covariate information providing the lowest overall errors. We also see individual cases where the stratified mean produced the lowest absolute error and instances where including covariates in VAST models actually increase the absolute error. When we reduce the number of strata that are included in indexing calculations to simulation species shifting into new territory, we typically see an increase in absolute error (as expected), though there are some scenarios where the impact is minimal.

(see Table \@ref(tab:CodERROR)) In the Cod results, when using reduced strata adding covariates produces worst VAST results (though still better than stratified mean). VAST without covariates much worse than stratified mean in fall with increasing temperature and all strata, but adding covariates corrects this.

(see Table \@ref(tab:HadERROR)) For Haddock, VAST has a particularly hard time in spring regularly producing larger errors than the stratified mean with added covariates only improving to the level of the stratified mean. VAST shows improved results in fall relative to the stratified mean with added covariates producing extremely low errors in some cases.

@@

<font size="5">*Cod Results*</font>

Of the scenarios without covariates, $12/16=75$\% had a VAST fit with a lower relative error compared to the stratified mean estimate. Including detailed covariate information in our Cod VAST estimates resulted in all $16/16=100\%$ having a lower relative error than the corresponding stratified mean estimate. The 4 scenarios without covariates that had a higher VAST error were each overestimates for the fall season that involved an increasing temperature gradient. As seen in Figure XXX, Cod are shifting their spatial distribution much more in the fall than they are in the spring. While adding covariates to these fits allowed for an improved error relative to the stratified mean, the VAST abundance still display an overestimate early in the time series before covariates correct the trend, allowing for a lower overall absolute error **(show these plots? IncTemp fall scenarios)**.

<font size="5">*Haddock Results*</font>

$10/16\approx63$\% of the scenarios without covariates had a VAST fit with a lower relative error compared to the stratified mean estimate. The 6 covariate-free VAST estimates that resulted in a higher error than the stratified mean spanned all scenarios and seasons, with several fall errors being especially large (significant overestimates). Adding covariate information resulted in $14/16\approx88$\% of VAST estimates having improved error compared to the stratified mean. The 2 scenarios that procedure worse error spanned temperature scenarios, but were both in the spring season when the proportion of the population in each strata remained constant in each scenario.**why?? doesnt make sense. the values with cov are about the same as SM** Qualitatively, including covariate information has a larger impact on decreasing the relative error compared to the stratified mean compared to failing to include covariates.

<font size="5">*Yellowtail Flounder Results*</font>

Of the scenarios without covariates, $33/48\approx68$\% had a VAST fit with a lower relative error compared to the stratified mean estimate. All 15 of the covariate-free VAST estimates that resulted in a higher error than the stratified mean were for the Fall season. These 15 Fall estimates span all other scenario variations and had a common theme of producing abundance estimates that are well above the true model value. Including detailed covariate information made a noticeable difference in our Yellowtail Flounder VAST estimates. Of the VAST estimates that included covariates information, $47/48\approx98$\% had a lower relative error than the corresponding stratified mean estimate. 
**For discussion: clearly our perfect covariate information made a big difference here, but the question is could we achieve similar results without "perfect" information? Which covariate is best? Good question for future research**

A more careful visual analysis of all error plots for each species reveals that VAST estimates tend to provide error estimates that are too high, while the stratified mean estimates are, on average, too low. Additionally, model/estimate ratio plots for all VAST estimates remain much closer to the desired value of 1 compared to stratified mean estimates, which can range as high as 10 and exhibit large yearly changes.

@@

```{r YTpct, echo=FALSE,out.width="95%",fig.show='hold', fig.align='center', fig.cap="add caption"}
knitr::include_graphics(c("Images/PercentPlots_ConPopConTemp_YT.png","Images/PercentPlots_ConPopIncTemp_YT.png"))
```

@@

tabl <- "
|	Parameter	|	Description | Settings A | Settings B
|:--	|:------- |:---- |:----
| *ObsModel* | Link function and assumed distribution | c(10,2) | c(10,2)
|FieldConfig | Specified spatial and/or spatio-temporal variation in predictors| c(Omega1=0, Epsilon1=0, Omega2=1, Epsilon2=1) | c(Omega1=0, Epsilon1=0, Omega2=1, Epsilon2=1)
| RhoConfig | Specifying whether intercepts or spatio-temporal variation is structured among time intervals | c(Beta1=3, Beta2=, Epsilon1=0, Epsilon2=4) | c(Beta1=3, Beta2=3, Epsilon1=0, Epsilon2=4)
| X1_formula | Right-sided formula affecting the 1st linear predictor  | X1_formula = ~ poly(Temp, degree=2 ) | N/A
| X2_formula | Right-sided formula affecting the 2nd linear predictor | X2_formula = ~ poly(Temp, degree=2 ) + poly(Habitat, degree=2 ) | X2_formula = ~ poly(Temp, degree=2 ) + poly(Habitat, degree=2 )

### 2023-03-03 Manuscript/Draft1.tex
manuscript update

We compare abundance estimates from stratified mean calculations to those derived from the Vector Autoregressive Spatio-Temporal (VAST) model. The stratified mean is a design-based approach that calculates the geometric mean catch per tow and has traditionally been used with stratified random sample designs. VAST is a spatial delta-generalized linear mixed model that estimates both abundance (biomass) and probability of occurrence (presence/absence) (Thorson 2019). If desired, VAST also allows users to include covariate data to better inform the model. Covariates can be static (eg. habitat preferences) or dynamics (eg. temperature). We explore whether including environmental predictors can help inform models and provide better abundance estimates, which is particularly relevant as climate change progresses. The stratified mean calculations are straightforward and quick, while VAST models require numerous user inputs and can take on the order of hours to complete. We compare the yearly estimated of abundance obtained from the stratified mean to estimates obtained from the Vector-Autoregressive Spatio-temporal (VAST) model.

We follow the advice given in (Thorson 2019) to build VAST models to estimate biomass on Georges Bank using stratified mean samples from our model output. In addition to exploring different link functions and assumed distributions, our VAST model-building process included testing the impact of including spatial and/or spatio-temporal variation in our models, considering varying number of knots in our mesh, and testing different forms of temporal correlation. We also carried out the same process both including covariates in our model as well as running models without covariate information. We considered covariates in the form of dynamic temperature values and/or static habitat values from our population model. When using covariates we ultimately decided to provide the most information to the model by including both covariates for both linear predictors. Since we know the true population values in our models we calculate the absolute error of each VAST estimate to compare between potential settings. Through this process, and in consultation with the VAST package creator, we determined setting that allowed VAST models to converge for all of our scenarios while also providing the lowest absolute error values. Settings for our VAST models can be seen in Table ???.

Our goal is to determine indexing approaches and settings that are robust to future environmental conditions and resulting spatial biomass patterns. An underlying assumption in all indexing methods is that individual random samples combine to accurately represent true abundance by a) containing a low enough noise level in the samples to allow for a discernible pattern and b) sampling all strata in which the population exists. These assumptions can be questioned given enough noise in the sampling process \textbf{cite?} and/or climate change causing a population to move into previously uninhabited strata. To simulate the impact of noise, we compare indexing estimates after adding noise to our samples versus those using the true sampling values. \textbf{BL: Help with correct notation for adding noise}. We simulate the effect of populations moving into new habitat by comparing indexing estimates using samples from all strata inhabited by each species on Georges bank {[}{[}Bank{]}{]} to those that only include a subset of the full spatial domain for each species. We chose strata to exclude for each species by reviewing how spatial preferences evolved in our increasing temperature scenarios and removing strata that each species either shifted into, or away from. Tabl XXX {[}{[}needs to be added{]}{]} lists all strata inhabited by each species, those that are removed from certain calculations, and the explanation of why these strata were removed.

\textbf{BL: I tried to provid emore detail above and also added another table}
{[}{[}may need to spell the aspect of reduced spatial domain a bit more because it might be confusing to readers who are expected us to just add areas around the current ones instead of reducing the strata{]}{]}

@@

The goal of our project was to analyze how well contemporary population indexing methods can track population trends under a variety of conditions, as depicted in Table \ref{tab:scenarios} Historically, Atlantic Cod has seen significant decline over the last XXX years while Haddock has increased in abundance in recent year {[}{[}can cite the 2022 management track assessments, see \url{https://apps-nefsc.fisheries.noaa.gov/saw/sari.php} for when the document becomes available{]}{]} \textbf{cite}. For this reason we compare indexing estimates using stratified random samples from decreasing population scenarios for Cod (see Table ???) and increasing population scenarios for Haddock (see Table ???). To provide a comprehensive analysis of population indexing methods we consider all possible scenario combinations for Yellowtail Flounder (see Table ???). The specific population trends used in our analyses can be see in Figure \ref{fig:pop-scenarios}

@@

General themes that exist in Tables X, Y and Z are that VAST estimates provide lower errors relative to those derived from the stratified mean, with VAST models that include covariate information providing the lowest overall errors. We also see individual cases where the stratified mean produced the lowest absolute error and instances where including covarites in VAST models actually increase the absolute error. When we reduce the number of strata that are included in indexing calculations to simulation species shifting into new territory, we typically see an increase in absolute error (as expected), though there are some scenarios where the impact is minimal.

In the Cod results, when using reduced strata adding covariates produces worst VAST results (though still better than stratified mean). VAST without covariates much worse than stratified mean in fall with increasing temperature and all strata, but adding covariates corrects this.

For Haddock, VAST has a particularly hard time in spring regularly producing larger errors than the stratified mean with added covariates only improving to the level of the stratified mean. VAST shows improved results in fall relative to the stratified mean with added covaraites producing extremely low errors in some cases.

@@

@@

@@

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.13}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.53}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.33}}@{}}
\toprule
\begin{minipage}[b]{\linewidth}\raggedright
Parameter
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Description
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Input/Value
\end{minipage} \\
\midrule
\endhead
\emph{ObsModel} & Link function and assumed distribution & c(10,2) \\
FieldCOnfig & Specified spatial and/or spatio-temporal variation in predictors & c(Omega1=0, Epsilon1=0, Omega2=1, Epsilon2=1) \\
RhoConfig & Specifying whether intercepts or spatio-temporal variation is structured among time intervals & c(Beta1=3, Beta2=3, Epsilon1=0, Epsilon2=4) \\
X1\_formula & Right-sided formula affecting the 1st linear predictor & X1\_formula = \textasciitilde{} poly(Temp, degree=2 ) \\
X2\_formula & Right-sided formula affecting the 2nd linear predictor & X2\_formula = \textasciitilde{} poly(Temp, degree=2 ) + poly(Habitat, degree=2 ) \\
\bottomrule
\end{longtable}

>>>>

We compare yearly abundance estimates obtained from the stratified mean to estimates obtained from the Vector-Autoregressive Spatio-temporal (VAST) model. The stratified mean is a design-based approach that calculates the geometric mean catch per tow and has traditionally been used with stratified random sample designs. VAST is a spatial delta-generalized linear mixed model that estimates both abundance (biomass) and probability of occurrence (presence/absence) (Thorson 2019). If desired, VAST also allows users to include covariate data to better inform the model. Covariates can be static (eg. habitat preferences) or dynamics (eg. temperature). We explore whether including environmental predictors can help inform models and provide better abundance estimates, which is particularly relevant as climate change progresses. The stratified mean calculations are straightforward and quick, while VAST models require numerous user inputs and can take on the order of hours to complete.

We follow the advice given in (Thorson 2019) to build VAST models to estimate biomass on Georges Bank using stratified random samples from our model output. In addition to exploring different link functions and assumed distributions, our VAST model-building process involved testing the impact of including spatial and/or spatio-temporal variation in our models, considering varying number of knots in our mesh, and testing different forms of temporal correlation. We also carried out the same process running models without covariate information as well as including covariates in our model. We considered covariates in the form of dynamic temperature values and/or static habitat values (\(Hab_{J.s}\)) from our population model. When using covariates we ultimately decided to provide the most information to the model by including both temperature and habitat covariates for both linear predictors. Knowing the true population values in our models allowed us to calculate the absolute error of each VAST estimate to compare between potential settings. Through this process, and in consultation with the VAST package creator, we ultimately compared the performance of two sets of settings in our VAST models, which can be seen in Table\ref{tab:VASTsettings}.

Our goal is to determine indexing approaches and settings that are robust to future environmental conditions and resulting spatial biomass patterns. An underlying assumption in all indexing methods is that individual random samples combine to accurately represent true abundance by a) containing a low enough noise level in the samples to allow for a discernible pattern and b) sampling all strata in which the population exists. These assumptions can be questioned given enough noise in the sampling process \textbf{cite?} and/or climate change causing a population to move into previously uninhabited strata. To simulate the impact of noise, we compare indexing estimates after adding noise to our samples versus those using the true sampling values. \textbf{BL: Help with correct notation for adding noise}. We simulate the effect of populations moving into new habitat by comparing indexing estimates using samples from all strata inhabited by each species on Georges Bank to those that only include a subset of the full spatial domain for each species. We chose strata to exclude for each species by reviewing how spatial preferences evolved in our increasing temperature scenarios and removing strata that each species either shifted into, or away from. Tabl XXX {[}{[}needs to be added{]}{]} lists all strata inhabited by each species, those that are removed from certain calculations, and the explanation of why these strata were removed.

@@

The goal of our project was to analyze the ability of contemporary population indexing methods to track population trends under a variety of conditions, as depicted in Table \ref{tab:scenarios}. Historically, Atlantic Cod has seen significant decline over the last XXX years while Haddock has increased in abundance in recent year {[}{[}can cite the 2022 management track assessments, see \url{https://apps-nefsc.fisheries.noaa.gov/saw/sari.php} for when the document becomes available{]}{]} \textbf{cite}. For this reason we compare indexing estimates using stratified random samples from decreasing population scenarios for Cod and increasing population scenarios for Haddock. To provide a comprehensive analysis of population indexing methods we consider all possible scenario combinations for Yellowtail Flounder. The specific population trends used in our analyses can be see in Figure \ref{fig:pop-scenarios}.

@@

Tables \ref{tab:YTresultsallstrata}, \ref{tab:YTresultsreducedstrata}, \ref{tab:CodERROR}, and \ref{tab:HadERROR} contain the results of our comparison between the absolute error of abundance estimates and model output. General themes are that VAST estimates provide lower errors relative to those derived from the stratified mean, with VAST models that include covariate information providing the lowest overall errors. We also see individual cases where the stratified mean produced the lowest absolute error and instances where including covariates in VAST models actually increase the absolute error. When we reduce the number of strata that are included in indexing calculations to simulation species shifting into new territory, we typically see an increase in absolute error (as expected), though there are some scenarios where the impact is minimal.

(see Table \ref{tab:CodERROR}) In the Cod results, when using reduced strata adding covariates produces worst VAST results (though still better than stratified mean). VAST without covariates much worse than stratified mean in fall with increasing temperature and all strata, but adding covariates corrects this.

(see Table \ref{tab:HadERROR}) For Haddock, VAST has a particularly hard time in spring regularly producing larger errors than the stratified mean with added covariates only improving to the level of the stratified mean. VAST shows improved results in fall relative to the stratified mean with added covariates producing extremely low errors in some cases.

@@

\emph{Cod Results}

Of the scenarios without covariates, \(12/16=75\)\% had a VAST fit with a lower relative error compared to the stratified mean estimate. Including detailed covariate information in our Cod VAST estimates resulted in all \(16/16=100\%\) having a lower relative error than the corresponding stratified mean estimate. The 4 scenarios without covariates that had a higher VAST error were each overestimates for the fall season that involved an increasing temperature gradient. As seen in Figure XXX, Cod are shifting their spatial distribution much more in the fall than they are in the spring. While adding covariates to these fits allowed for an improved error relative to the stratified mean, the VAST abundance still display an overestimate early in the time series before covariates correct the trend, allowing for a lower overall absolute error \textbf{(show these plots? IncTemp fall scenarios)}.

\emph{Haddock Results}

\(10/16\approx63\)\% of the scenarios without covariates had a VAST fit with a lower relative error compared to the stratified mean estimate. The 6 covariate-free VAST estimates that resulted in a higher error than the stratified mean spanned all scenarios and seasons, with several fall errors being especially large (significant overestimates). Adding covariate information resulted in \(14/16\approx88\)\% of VAST estimates having improved error compared to the stratified mean. The 2 scenarios that procedure worse error spanned temperature scenarios, but were both in the spring season when the proportion of the population in each strata remained constant in each scenario.\textbf{why?? doesnt make sense. the values with cov are about the same as SM} Qualitatively, including covariate information has a larger impact on decreasing the relative error compared to the stratified mean compared to failing to include covariates.

\emph{Yellowtail Flounder Results}

Of the scenarios without covariates, \(33/48\approx68\)\% had a VAST fit with a lower relative error compared to the stratified mean estimate. All 15 of the covariate-free VAST estimates that resulted in a higher error than the stratified mean were for the Fall season. These 15 Fall estimates span all other scenario variations and had a common theme of producing abundance estimates that are well above the true model value. Including detailed covariate information made a noticeable difference in our Yellowtail Flounder VAST estimates. Of the VAST estimates that included covariates information, \(47/48\approx98\)\% had a lower relative error than the corresponding stratified mean estimate.
\textbf{For discussion: clearly our perfect covariate information made a big difference here, but the question is could we achieve similar results without ``perfect'' information? Which covariate is best? Good question for future research}

A more careful visual analysis of all error plots for each species reveals that VAST estimates tend to provide error estimates that are too high, while the stratified mean estimates are, on average, too low. Additionally, model/estimate ratio plots for all VAST estimates remain much closer to the desired value of 1 compared to stratified mean estimates, which can range as high as 10 and exhibit large yearly changes.

@@

{\centering \includegraphics[width=0.95\linewidth]{Images/PercentPlots_ConPopConTemp_YT} \includegraphics[width=0.95\linewidth]{Images/PercentPlots_ConPopIncTemp_YT}

\caption{add caption}\label{fig:YTpct}
\end{figure}

@@

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.14}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.38}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.24}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.24}}@{}}
\toprule
\begin{minipage}[b]{\linewidth}\raggedright
Parameter
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Description
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Settings A
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Settings B
\end{minipage} \\
\midrule
\endhead
\emph{ObsModel} & Link function and assumed distribution & c(10,2) & c(10,2) \\
FieldConfig & Specified spatial and/or spatio-temporal variation in predictors & c(Omega1=0, Epsilon1=0, Omega2=1, Epsilon2=1) & c(Omega1=0, Epsilon1=0, Omega2=1, Epsilon2=1) \\
RhoConfig & Specifying whether intercepts or spatio-temporal variation is structured among time intervals & c(Beta1=3, Beta2=, Epsilon1=0, Epsilon2=4) & c(Beta1=3, Beta2=3, Epsilon1=0, Epsilon2=4) \\
X1\_formula & Right-sided formula affecting the 1st linear predictor & X1\_formula = \textasciitilde{} poly(Temp, degree=2 ) & N/A \\
X2\_formula & Right-sided formula affecting the 2nd linear predictor & X2\_formula = \textasciitilde{} poly(Temp, degree=2 ) + poly(Habitat, degree=2 ) & X2\_formula = \textasciitilde{} poly(Temp, degree=2 ) + poly(Habitat, degree=2 ) \\
\bottomrule
\end{longtable}
